New insights into the Weddell Sea ecosystem applying a quantitative network approach

### 2023-11-10 Manuscript/Manuscript.tex
Updated biblio Pecuchet

% Advances in Geosciences (adgeo)
% Advances in Radio Science (ars)
% Advances in Science and Research (asr)
% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
% Annales Geophysicae (angeo)
% Archives Animal Breeding (aab)
% Atmospheric Chemistry and Physics (acp)
% Atmospheric Measurement Techniques (amt)
% Biogeosciences (bg)
% Climate of the Past (cp)
% DEUQUA Special Publications (deuquasp)
% Drinking Water Engineering and Science (dwes)
% Earth Surface Dynamics (esurf)
% Earth System Dynamics (esd)
% Earth System Science Data (essd)
% E&G Quaternary Science Journal (egqsj)
% EGUsphere (egusphere) | This is only for EGUsphere preprints submitted without relation to an EGU journal.
% European Journal of Mineralogy (ejm)
% Fossil Record (fr)
% Geochronology (gchron)
% Geographica Helvetica (gh)
% Geoscience Communication (gc)
% Geoscientific Instrumentation, Methods and Data Systems (gi)
% Geoscientific Model Development (gmd)
% History of Geo- and Space Sciences (hgss)
% Hydrology and Earth System Sciences (hess)
% Journal of Bone and Joint Infection (jbji)
% Journal of Micropalaeontology (jm)
% Journal of Sensors and Sensor Systems (jsss)
% Magnetic Resonance (mr)
% Mechanical Sciences (ms)
% Natural Hazards and Earth System Sciences (nhess)
% Nonlinear Processes in Geophysics (npg)
% Ocean Science (os)
% Polarforschung - Journal of the German Society for Polar Research (polf)
% Primate Biology (pb)
% Proceedings of the International Association of Hydrological Sciences (piahs)
% Safety of Nuclear Waste Disposal (sand)
% Scientific Drilling (sd)
% SOIL (soil)
% Solid Earth (se)
% The Cryosphere (tc)
% Weather and Climate Dynamics (wcd)
% Web Ecology (we)
% Wind Energy Science (wes)

@@

\begin{abstract}
Network approaches can shed light on the structure and stability of
complex marine communities. In recent years, such approaches have been
successfully applied to study polar ecosystems, improving our knowledge
on how they might respond to ongoing environmental changes. The Weddell
Sea is one of the most studied marine ecosystems outside the Antarctic
Peninsula in the Southern Ocean. Yet, few studies consider the known
complexity of the Weddell Sea food web, which in its current form
comprises 490 species and 16041 predator-prey interactions. Here we
analysed the Weddell Sea food web, focusing on the species and trophic
interactions that underpin ecosystem structure and stability. We
estimated the strength for each interaction in the food web,
characterised species position in the food web using unweighted and
weighted food web properties, and analysed species' roles with respect
to the stability of the food web. We found that the distribution of the
interaction strength at the food web level is asymmetric, with many weak
interactions and few strong ones. We detected a positive relationship
between species mean interaction strength and two unweighted properties
(i.e., trophic level and the total number of interactions). We also
found that only a few species possess key positions in terms of food web
stability. These species are characterised by high mean interaction
strength, mid to high trophic level, relatively high number of
interactions, and mid to low trophic similarity. In this study, we
integrated unweighted and weighted food web information, enabling a more
complete assessment of the ecosystem structure and function of the
Weddell Sea food web. Our results provide new insights, which are
important for the development of effective policies and management
strategies, particularly given the ongoing initiative to implement a
Marine Protected Area (MPA) in the Weddell Sea.
\end{abstract}

\section{Introduction}

Food web analysis constitutes an important framework for understanding
ecological community structure and for conserving biodiversity through
ecosystem management \citep{Thompson2012}. Although topological food web
analysis, which considers only the presence and absence of predator-prey
interactions, provides important insights into the structure and
functioning of ecological communities
\citep[e.g.][]{Pascual2006, Kortsch2015, Marina2018, Cordone2020, Rodriguez2022},
more information on the nature of the trophic interactions is needed to
effectively charaterise ecosystem dynamics and stability
\citep{Kortsch2021}. This is a fundamental step for providing
assemessments on ecosystem vulnerability to environmental pressures and
for prioritising management actions. In this regard, quantifying the
strength of trohic interactions and species' roles within the network
are of paramount importance
\citep{Carrara2015, Allesina2015, Nilsson2016, Cirtwill2018a}.

Estimating interaction strength in food webs allows differentiating the
importance of species interactions. On the contrary, unweighetd food web
representations give equal importantance to all interactions
well-knowing that some species interactions are stronger than others and
hence play a diffrent role for ecosystem functioning and stability. Both
empirical and theoretical studies show that interactions strength
distributions in food webs are assymetric
\citep{Paine1992, McCann1998, Emmerson2004, Wootton2005, Kortsch2021},
containing a few strong and many weak interactions. This asymmetric
patterning of weak and strong links is crucial to food web stability
\citep{Paine1992, McCann1998, Neutel2002}. In a recent paper on an
aquatic food web it was further highlighted that temporal changes in
ecosytem functioning could only be predicted using weighted food web
structure \citep{Kortsch2021}. Hence, in order to assess the stability
and functioning of a food web, it is important to first determine the
quantitative structure of the trophic network.

Several methodologies have been applied to estimate interaction strength
in food webs, where the quantity and quality of the data mostly
determines which approach is the most convenient \citep{Berlow2004}.
Approaches include experimental methods combined with dynamic modelling
\citep{Emmerson2004, Carrara2015}, measurements of species abundances
through time \citep{Fahimipour2014, Chang2021}, and estimation of
metabolic rates and biomass of all species in the community
\citep{Neutel2014}. However, these types of methods require large
experimental set-ups and parameterisations restricting the analyses to
smaller networks (e.g., approximatley 10 species or less). Other methods
based on allometric scaling relationships and biomass information
\citep{Kortsch2021, Gauzens2019} can be applied to larger networks with
less data requirements, but this comes at the expense of precision in
the predictions. For even larger food webs composed of nearly 1000
species and more than 10000 interactions, only methods with even less
data requirements are feasible. One of these methods, proposed by
\citet{Pawar2012}, combines data on consumer and resource body masses
and consumer search space (interaction dimensionality) to obtain
interaction strength estimates for each pairwise predator-prey
interaction. An advantage of this method is that it can be applied
without information on species biomass.

Using a a network approach, different types (e.g., terrestial, lake,
marine) of food webs from various geographic locations are studied
worldwide, including marine polar food webs
\citep{Carscallen2012, Santana2013, Kortsch2019, Pecuchet2022}. Some of
the studies from the Arctic show how food web properties (e.g.,
connectance) are constrained by environmental factors such as sea ice
cover and seawater temperature \citep{Kortsch2019, Pecuchet2022}. In the
Southern Ocean, important insights have been gained into mechanisms of
energy flow, the relative importance of individual species and their
traits, and the influence of environmental variables (e.g., sea-ice) on
the structure of local food webs \citep{Cordone2020, Rossi2019}. For
instance, in Potter Cove (West Antarctic Peninsula) the substratum type
(i.e.~hard/soft or rocks/sediments) plays a significant role in the
structure and stability of the food web. In Terra Nova Bay (Ross Sea),
the architecture of biodiversity was reshaped by the pulsed input of
sympagic food sources following sea-ice break up, with food web
simplification, decreased intraguild predation, potential disturbance
propagation and increased vulnerability to biodiversity loss
\citep{Rossi2019}.

The Weddell Sea is expected to be one of the last regions of the
Southern Ocean to experience the consequences of climate change due to
its extensive ice cover and ocean currents \citep{Teschke2021} resulting
in less sea surface warming compared to other areas of the Southern
Ocean. This Sea plays an important role in driving global thermohaline
circulation and ventilating the global abyssal ocean because it
generates a considerable part of the Antarctic Bottom Water
\citep{Fahrbach2009}. Because of these environmental characteristics,
the Weddell Sea may serve as a refuge for Antarctic species which depend
on sea ice (e.g.~krill, emperor penguin, Weddell seal) or have low heat
tolerance (e.g., most notothenioid fishes) due to their adaptations to
freezing tempeartures \citep{Griffiths2017}. While essential large-scale
hydrodynamic relationships are relatively well-known for this region
\citep{deSteur2019}, information on the current distribution, abundance
and sensitivity to climate change is only partially known for a few
species (e.g., emperor penguin) \citep{Houstin2022}.

The network complexity of the Weddell Sea food web is high comprising
488 species and 16200 predator-prey interactions \citep{Jacob2011}. In
an attempt to better understand species roles related to food web
stability, Jacob et al.~(2011) performed secondary extinction
experiments and found that the removal of small to medium-sized, and not
large, organisms caused a cascade of secondary extinctions. This
findings highlighted the relative importance of predators, rather than
prey, for the architecture, functioning and stability of the Weddell Sea
food web, which coincides with findings from recent meta-analyses in
natural complex food webs \citep{Brose2019, Perkins2022}. Other
investigations considered this food web in a meta-analysis context
showing that high predator-prey body-mass ratios are found for predator
groups with specific trait combinations, including small vertebrates and
large swimming or flying predators \citep{Brose2019}. These trait
combinations generate weak interactions that stabilize communities
against perturbations maintaining ecosystem functioning.

In this study, we aim to go beyond a purely topological
(presence/absence) assessment of who eats whom in the Weddell Sea
ecosystem by providing a quantitative analysis of the trophic
interaction network. We aim to analyse the species' role for the
structure and stability of the food web. To achieve this, we: 1)
estimated the strength for each interaction in the Weddell Sea food web,
2) characterised species' role considering both weighted and unweighted
properties, and 3) analysed the species' role related to the stability
of the food web. This is the first time that interaction strengths were
estimated for all pairwise trophic interactions at the species level
(except for a few) for the Weddell sea food web.

@@

The high Antarctic Weddell Sea shelf is situated between 74 and 78ºS,
stretching approximately 450 km from East to West (Figure 1). Water
depth varies between 200 and 500 meters, and shallower areas are covered
by continental ice, which forms the coastline along the eastern and
southern part of the Weddell Sea. The shelf area contains a complex
three-dimensional benthic habitat with large benthic biomasses,
intermediate to high diversity in comparison to benthic boreal
communities and a spatially patchy distribution of organisms
\citep{Dayton1990, Teixido2002}.

@@

The Weddell Sea food web was retrieved from the GlobAL daTabasE of
traits and food Web Architecture (GATEWAy, version 1.0) of the German
Centre for Integrative Biodiversity Research (iDiv) Halle-Jena-Leipzig
\citep{Brose2018}. In addition to predator-prey interactions, the
database contains information on other biological data such as the mean
body mass and movement type for each species in the food web.
Furthermore, it incorporates information about the interaction itself,
such as the dimension of the predator search space (2 or 3 dimensions).
In its current form, the Weddell Sea food web comprises 490 species and
16041 predator-prey interactions and constitutes one of the most
resolved food webs constructed to date \citep{Jacob2011}.

@@

To estimate the strength of each pairwise interaction in the food web,
we followed an approach proposed by \citet{Pawar2012}. The minimum data
requirements are body mass of the consumer (predator) and resource
(prey), and the interaction dimensionality (ID) classified as 2 or 3
dimensions. The ID is defined as the search space dimensions of the
predator, which is also equivalent to the movement space of the prey.
Thus, the ID is classified as 2D when both predator and prey move in 2D
(e.g., both are benthic) or if a predator moves in 3D and a prey in 2D
(e.g., pelagic predator on benthic prey). The ID is classified as 3D
when both predator and prey move in 3D (e.g., both pelagic) or if the
predator moves in 2D and the prey in 3D (e.g., benthic predator, pelagic
prey) \citep{Pawar2012}. GATEWAy v.1.0 provides information on the mean
body mass for consumers and resources, except for `detritus' and
`sediment', and the dimensionality for the majority of the interactions,
though the latter is missing in some cases (924 interactions). To
complete the missing data on species `dimensionality', we used
information about the movement type of predators and prey included in
GATEWAy.

@@

We obtained estimates for resource density and the search rate from the
scaling relationships with the resource and the consumer mass,
respectively \citep{Pawar2012}. The coefficients of such relationships,
determined by ordinary least squares regression, vary with the
interaction dimensionality. On one hand, resource density scales with
resource mass as power-law with different exponents in 2D and in 3D.
Since mean mass for resources `phytodetritus' and `sediment' were not
available in GATEWAy, we considered the body mass of the smallest
phytoplankton species (`Fragilariopsis cylindrus') as a proxy. This is
justified by the fact that `phytodetritus' and `sediment' are mainly
composed of dead or senescent phytoplankton reaching the seabed
\citep{Wolanski2011}. On the other hand, search rate scales with
consumer mass as power-law with exponents in 2D and in 3D.

We fitted six candidate models (Exponential, Gamma, log-Normal, Normal,
Power-law and Uniform) the interaction strength distribution using
maximum likelihood \citep{McCallum2008}, and selected the best fitting
model by computing the Akaike Information Criterion AIC
\citep{Burnham2002}.

@@

To characterise the role of each species in the food web, we considered
unweighted and weighted food web properties (Figure 2). Unweighted
properties are related to properties commonly used in qualitative food
web studies and only describe the presence or absence of interactions
without any information on strength between a pairwise species link
\citep{Martinez1991, Dunne2002, Borrelli2014}. In contrast, weighted
properties capture the importance of a trophic interaction by
considering its strength.

To assess species roles as a function of the weighted food web, we
focused on mean interaction strength defined as the average strength of
all interactions for a given species. Further we calculated three
unweighted species properties: a) species degree, i.e., the sum of in-
and out-going interactions ; b) trophic level ; and c) trophic
similarity, i.e., the trophic overlap based on shared and unique
resources and consumers. These metrics were chosen to assess a species
role based on the unweighted network. The species degree has often been
equated with species importance to the structure and functioning within
a food web, i.e.~perturbations to high-degree species may therefore have
more significant effects on the food web robustness to perturbations
than low-degree species
\citetext{\citealp{Dunne2002a}; \citealp[references
in][]{Cirtwill2018a}}. The trophic level offers information about how
important a species is to its biotic community, i.e., top predators and
primary producers are expected to have particularly large effects on the
rest of their communities through top-down and bottom-up control,
respectively \citep[references in][]{Cirtwill2018a}. Trophic similarity
is an index of trophic overlap considering the set of prey and predators
for a pair of species; it measures one of the most important aspects of
species' niches, the trophic niche, and functional aspects of
biodiversity \citep{Martinez1991, Williams2000}.

Furthermore, we took species habitat affiliation into account, which
describes the physical position of a species within the ecosystem.
Species were categorised as: 1) benthic, if a species lives on the
seafloor; 2) pelagic, if a species lives close to the surface; 3)
benthopelagic, if it moves between and connects the aforementioned
environments; 4) demersal, if it lives and feeds on, or near, the bottom
of the sea; and 5) land-based, if the consumer is not strictly aquatic
but feeds predominantly on marine species. Species habitat affiliations
were retrieved from \citet{Jacob2011}.

To study the relationship between species mean interaction strength
(weighted property) and the unweighted species properties, we performed
linear regression analyses between the log mean interaction strength and
each of the aforementioned unweighted properties. Thus, we considered
the interaction strength as the dependent variable and the given
unweighted property as the independent variable, and obtained the
coefficients (slope and intercept) for the linear model. Models were
fitted using the least squares approach. We also explored the mean
interaction strength distribution with the species habitat.

@@

To analyse the impact of species on food web stability, we performed
extinction simulations deleting one species at a time, that is for every
extinction, network size was reduced by one species only. After each
extinction, we calculated the stability of the network minus the removed
species (489 nodes) and compared it with that of the whole network (490
nodes in total). To calculate stability, we used the mean of the real
part of the maximum eigenvalue of the Jacobian matrix using randomized
Jacobians, keeping the predator-prey sign structure fixed
\citep{Allesina2008, Grilli2016}. This stability index indicates a more
stable food web when it is negative. We performed 1000 simulations for
each species removal and obtained a mean maximum eigenvalue for each
case. Finally, we statistically analysed this difference with an
Anderson-Darling test considering a p-value \textless{} 0.01
\citep{Scholz1987}. If the difference between the networks without and
with the removed species is positive, then the stability of the whole
food web is higher without the targeted species, in other words this
species makes the network less stable. If the difference is negative,
then the stability of the whole food web is lower without the targeted
species, i.e., this species ahs a stabilizing effect. A detailed
description on the stability calculations can be found in the
supplementary material.

To identify the species with the highest effect on food web stability,
we plotted the results of each species' extinction and its effect on
food web stability, considering weighted (interaction strength) and
unweighted properties, and species habitat.

@@

The statistical distribution that best fitted the empirical interaction
strength distribution of the Weddell Sea food web was a `gamma' due to
the high proportion of weak interactions and the existence of a few
strong interactions (Figure 3, Table S3).

>>>>

% Advances in Geosciences (adgeo)
% Advances in Radio Science (ars)
% Advances in Science and Research (asr)
% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
% Aerosol Research (ar)
% Annales Geophysicae (angeo)
% Archives Animal Breeding (aab)
% Atmospheric Chemistry and Physics (acp)
% Atmospheric Measurement Techniques (amt)
% Biogeosciences (bg)
% Climate of the Past (cp)
% DEUQUA Special Publications (deuquasp)
% Earth Surface Dynamics (esurf)
% Earth System Dynamics (esd)
% Earth System Science Data (essd)
% E&G Quaternary Science Journal (egqsj)
% EGUsphere (egusphere) | This is only for EGUsphere preprints submitted without relation to an EGU journal.
% European Journal of Mineralogy (ejm)
% Fossil Record (fr)
% Geochronology (gchron)
% Geographica Helvetica (gh)
% Geoscience Communication (gc)
% Geoscientific Instrumentation, Methods and Data Systems (gi)
% Geoscientific Model Development (gmd)
% History of Geo- and Space Sciences (hgss)
% Hydrology and Earth System Sciences (hess)
% Journal of Bone and Joint Infection (jbji)
% Journal of Micropalaeontology (jm)
% Journal of Sensors and Sensor Systems (jsss)
% Magnetic Resonance (mr)
% Mechanical Sciences (ms)
% Natural Hazards and Earth System Sciences (nhess)
% Nonlinear Processes in Geophysics (npg)
% Ocean Science (os)
% Polarforschung - Journal of the German Society for Polar Research (polf)
% Primate Biology (pb)
% Proceedings of the International Association of Hydrological Sciences (piahs)
% Safety of Nuclear Waste Disposal (sand)
% Scientific Drilling (sd)
% SOIL (soil)
% Solid Earth (se)
% State of the Planet (sp)
% The Cryosphere (tc)
% Weather and Climate Dynamics (wcd)
% Web Ecology (we)
% Wind Energy Science (wes)

@@

\begin{abstract}
Network approaches can shed light on the structure and stability of
complex marine communities. In recent years, such approaches have been
successfully applied to study polar ecosystems, improving our knowledge
on how they might respond to ongoing environmental changes. The Weddell
Sea is one of the most studied marine ecosystems outside the Antarctic
Peninsula in the Southern Ocean. Yet, few studies consider the known
complexity of the Weddell Sea food web, which in its current form
comprises 490 species and 16041 predator-prey interactions. Here we
analysed the Weddell Sea food web, focusing on the species and trophic
interactions that underpin ecosystem structure and stability. We
estimated the strength for each interaction in the food web,
characterised species position in the food web using unweighted and
weighted food web properties, and analysed species' roles with respect
to the stability of the food web. We found that the distribution of the
interaction strength (IS) at the food web level is asymmetric, with many
weak interactions and few strong ones. We detected a positive
relationship between species median IS and two unweighted properties
(i.e., trophic level and the total number of interactions). We also
found that only a few species possess key positions in terms of food web
stability. These species are characterised by high median IS, mid to
high trophic level, relatively high number of interactions, and mid to
low trophic similarity. In this study, we integrated unweighted and
weighted food web information, enabling a more complete assessment of
the ecosystem structure and function of the Weddell Sea food web. Our
results provide new insights, which are important for the development of
effective policies and management strategies, particularly given the
ongoing initiative to implement a Marine Protected Area (MPA) in the
Weddell Sea.
\end{abstract}

\introduction[Introduction]

Food web analysis constitutes an important framework for understanding
ecological community structure and for conserving biodiversity through
ecosystem management \citep{Thompson2012}. Although topological food web
analysis, which considers only the presence and absence of predator-prey
interactions, provides important insights into the structure and
functioning of ecological communities
\citep[e.g.][]{Pascual2006, Kortsch2015, Marina2018, Cordone2020, Rodriguez2022},
more information on the nature of the trophic interactions is needed to
effectively characterise ecosystem dynamics and stability
\citep[e.g.][]{Kortsch2021, Pecuchet2022}. This is a fundamental step
for providing assessments on ecosystem vulnerability to environmental
pressures and for prioritising management actions. In this regard,
quantifying the strength of trophic interactions and species' roles
within the network are of paramount importance
\citep{Carrara2015, Allesina2015, Nilsson2016, Cirtwill2018a}.

Estimating interaction strength (IS) in food webs allows differentiating
the importance of species interactions. On the contrary, unweighted food
web representations give equal importance to all interactions
well-knowing that some species interactions are stronger than others and
hence play a different role for ecosystem functioning and stability.
Both empirical and theoretical studies show that interactions strength
distributions in food webs are asymmetric
\citep{Paine1992, McCann1998, Emmerson2004, Wootton2005, Kortsch2021},
containing a few strong and many weak interactions. This asymmetric
patterning of weak and strong links is crucial to food web stability
\citep{Paine1992, McCann1998, Neutel2002}. In a recent paper on an
aquatic food web it was further highlighted that temporal changes in
ecosystem functioning could only be predicted using a weighted food web
structure \citep{Kortsch2021}. Hence, in order to assess the stability
and functioning of a food web, it is important to first determine link
weights.

Several methodologies have been applied to estimate IS in food webs,
where the quantity and quality of the data mostly determine which
approach is the most convenient \citep{Berlow2004}. Approaches include
experimental methods combined with dynamic modelling
\citep{Emmerson2004, Carrara2015}, measurements of species abundances
through time \citep{Fahimipour2014, Chang2021}, and estimation of
metabolic rates and biomass of all species in the community
\citep{Neutel2014}. However, these types of methods require large
experimental set-ups and parameterisations restricting the analyses to
smaller networks (e.g., approximately 10 species or less). Other methods
based on allometric scaling relationships and biomass information
\citep{Kortsch2021, Gauzens2019} can be applied to larger networks with
less data requirements, but this comes at the expense of precision in
the predictions. For even larger food webs composed of nearly 1000
species and more than 10000 interactions, only methods with even less
data requirements are feasible. One of these methods, proposed by
\citet{Pawar2012}, combines data on consumer and resource body masses
and consumer search space (interaction dimensionality) to obtain IS
estimates for each pairwise predator-prey interaction. An advantage of
this method is that it can be applied without information on species
biomass.

Diverse food webs (e.g., terrestrial, lake, marine) from various
geographic locations have been studied worldwide, including marine food
webs in polar regions
\citep{Carscallen2012, Santana2013, Kortsch2019, Pecuchet2022}. Studies
from the Arctic show that marine food web properties (e.g., connectance
and modularity) are constrained by sea ice and seawater temperature
\citep{Kortsch2019, Pecuchet2022}. Likewise for the Southern Ocean it
was shown that sea ice has impacts on local (Ross Sea) food web
structure \citep{Rossi2019}. For example, in the Ross Sea, sea-ice break
up introduced a surge of sympagic food, which led to simpler food webs
with reduced intraguild predation, potential disturbance propagation and
increased vulnerability to biodiversity loss \citep{Rossi2019}.

The Weddell Sea is expected to be one of the last regions of the
Southern Ocean to experience the consequences of climate change due to
its extensive sea ice cover and ocean currents \citep{Teschke2021}
resulting in less sea surface warming compared to other areas of the
Southern Ocean. Generaly, the Southern Ocean plays an important role in
driving global thermohaline circulation and ventilating the global
abyssal ocean because it generates a considerable part of the Antarctic
Bottom Water \citep{Fahrbach2009}. Because of these environmental
characteristics, the Weddell Sea may serve as a refuge for Antarctic
species which depend on sea ice (e.g.~krill, emperor penguin, Weddell
seal) or have low heat tolerance (e.g., most notothenoid fishes) due to
their adaptations to freezing temperatures \citep{Griffiths2017}. While
essential large-scale hydrodynamic relationships are relatively
well-known for this region \citep{deSteur2019}, information on the
current distribution, abundance and sensitivity to climate change is
only partially known for a few species (e.g., emperor penguin)
\citep{Houstin2022}.

The Weddell Sea food web is highly complex, comprising 488 species and
16200 predator-prey interactions \citep{Jacob2011}. To better understand
species roles related to food web stability, Jacob et al.~(2011)
performed secondary extinctions and found that the removal of small to
medium-sized, and not large, organisms caused a cascade of secondary
extinctions. These findings highlighted the relative importance of
predators, rather than prey, for the architecture, functioning and
stability of the Weddell Sea food web, which coincides with findings
from recent meta-analyses in empirical complex food webs
\citep{Brose2019, Perkins2022}. \citet{Brose2019} considered this food
web in a meta-analysis context showing that high predator-prey body-mass
ratios are found for predator groups with specific trait combinations,
including small vertebrates and large swimming or flying predators.
These trait combinations generate weak interactions that stabilize
communities against perturbations maintaining ecosystem functioning.

In this study, we aim to go beyond a purely topological
(presence/absence) assessment of who eats whom in the Weddell Sea
ecosystem by providing a quantitative analysis of the trophic network
structure. We aim to analyse the species' role for the structure and
stability of the food web. To achieve this, we: 1) estimated the
strength for each interaction, 2) characterised species' role
considering both weighted and unweighted properties, and 3) analysed
species' roles related to the stability of the food web.

@@

The high Antarctic Weddell Sea shelf is situated between 74 and 78ºS,
stretching approximately 450 km from East to West (Figure 1). Water
depth varies between 200 and 500 meters, and shallower areas are covered
by continental ice, which forms the coastline along the eastern and
southern parts of the Weddell Sea. The shelf area contains a complex
three-dimensional benthic habitat with large benthic biomasses,
intermediate to high diversity in comparison to benthic boreal
communities and a spatially patchy distribution of organisms
\citep{Dayton1990, Teixido2002}.

@@

The Weddell Sea food web was retrieved from the GlobAL daTabasE of
traits and food Web Architecture (GATEWAy, version 1.0) of the German
Centre for Integrative Biodiversity Research (iDiv) Halle-Jena-Leipzig
\citep{Brose2018}. In addition to predator-prey interactions, the
database contains information on other biological data such as mean body
mass and movement type for each species in the food web. Furthermore, it
incorporates information about the interaction itself, such as the
dimension of the predator search space (2 or 3 dimensions). In its
current form, the Weddell Sea food web comprises 490 species and 16041
predator-prey interactions and thus constitutes one of the most resolved
food webs constructed to date \citep{Jacob2011}.

Complex empirical food webs are compilations of species and their
potential feeding interactions within a given area and time period. The
species occurrence data to construct the Weddell Sea food web
(Antarctica) was sampled between 1983 and 2005, and hence the Weddell
Sea food web is representative for this time period. The diet
composition for each species was observed between 2001 and 2004 through
a combination of field observations and stomach content analyses
(further details can be found in \citet{Jacob2005}). Some species, such
as benthic grazers and suspension feeders had poor taxonomic resolution
of their prey, therefore laboratory information about their size,
behavior, and stable isotope signatures was used to infer their feeding
habits \citep{Jacob2011}. As the data used to construct the Weddell Sea
food web is sampled during the summer season due to the inaccessibility
of ecological field sampling in Antarctica during the winter, the
Weddell Sea food web is representative of the summer season. Moreover,
the species in the food web are representative of adult specimens, and
were not categorized into ontogenetic life stages (i.e., larvae,
juveniles, and adults). As a result, ontogenetic diet shifts cannot be
addressed with this food web. The average body masses of species were
either directly measured \citep{Jacob2005} or taken from published
accounts for marine mammals and seabirds \citep{Brose2006a}. Due to the
static nature of the Weddell Sea food web and its low temporal
resolution, it is not possible to assess phenological mismatches nor
species temporal turnover. Because of these methodological limitations,
the nature of the Weddell Sea food web data may likely lead to
overestimations of some trophic interactions and network properties
(connectance), compared to realised networks at any given point in time
and space. The spatial extent of the food web is described in the Study
area section (Figure 1).

@@

To estimate the strength of each pairwise interaction in the food web,
we followed an approach proposed by \citet{Pawar2012}. The minimum data
requirements are body mass of the consumer (predator) and resource
(prey), and the interaction dimensionality (ID) classified as 2 or 3
dimensions. The ID is defined as the search space dimensions of the
predator, which is also equivalent to the movement space of the prey.
Thus, the ID is classified as 2D when both predator and prey move in 2D
(e.g., both are benthic) or if a predator moves in 3D and a prey in 2D
(e.g., pelagic predator on benthic prey). The ID is classified as 3D
when both predator and prey move in 3D (e.g., both pelagic) or if the
predator moves in 2D and the prey in 3D (e.g., benthic predator, pelagic
prey) \citep{Pawar2012}.

GATEWAy v.1.0 provides information on the mean body mass for consumers
and resources, except for `detritus' and `sediment', and the
dimensionality for the majority of the interactions, though the latter
is missing in some cases (924 interactions). To complete the missing
data on species `dimensionality', we used information about the movement
type of predators and prey included in GATEWAy.

@@

We obtained estimates for resource density and the search rate from the
scaling relationships with the resource and the consumer mass,
respectively \citep{Pawar2012}. The coefficients of such relationships,
determined by ordinary least squares regression, vary with the ID. On
one hand, resource density scales with resource mass as power-law with
different exponents in 2D and in 3D. Since mean mass for resources
`phytodetritus' and `sediment' were not available in GATEWAy, we
considered the body mass of the smallest phytoplankton species
(`Fragilariopsis cylindrus') as a proxy. This is justified by the fact
that `phytodetritus' and `sediment' are mainly composed of dead or
senescent phytoplankton reaching the seabed \citep{Wolanski2011}. On the
other hand, search rate scales with consumer mass as power-law with
exponents in 2D and in 3D \citep{Pawar2012}. The equations for
estimating the search rate and the resource density are specified in the
Supplementary Material.

To quantify the variability in IS estimates, we considered the
uncertainties inherent in estimating the exponents via standard linear
regression. In this regard, the exponents reported in \citet{Pawar2012}
can be viewed as the mean of a normal random variable, with the standard
deviation given by the associated standard errors. We generated 1000
random samples and calculated an IS value for each sample, with the
final outcome of a distribution of IS estimates for each interaction.
Given that the IS distributions are right-skewed and bounded below zero,
we characterized the central tendency using the median IS. To represent
the distribution variability, we reported the interquartile range
(Supplementary Material, Figure S1). This approach accounts for the
propagation of uncertainty from the exponent estimates into the derived
IS values, allowing us to systematically assess the robustness of
conclusions based on these interaction metrics.

We fitted the IS distribution (i.e.~medians for each interaction) to six
candidate models (Exponential, Gamma, log-Normal, Normal, Power-law and
Uniform), using maximum likelihood \citep{McCallum2008}, and selected
the best fitting model by computing the Akaike Information Criterion AIC
\citep{Burnham2002}.

@@

To characterise the role of each species in the food web, we considered
unweighted and weighted food web properties (Figure 2). Unweighted
properties are related to properties commonly used in qualitative food
web studies and only describe the presence or absence of interactions
without any information on link strength
\citep{Martinez1991, Dunne2002, Borrelli2014}. In contrast, weighted
properties capture the importance of a trophic interaction by
considering its strength.

To assess species roles as a function of the weighted food web, we
focused on mean IS defined as the average strength of all interactions
for a given species. Further, we calculated three unweighted species
properties: a) species degree, i.e., the sum of in- and out-going
interactions ; b) trophic level ; and c) trophic similarity, i.e., the
trophic overlap based on shared and unique resources and consumers.
These metrics were chosen to assess species' roles based on the
unweighted network. The species degree has often been linked to a
species importance for the structure and functioning within a food web,
i.e.~perturbations to high-degree species may therefore have more
significant effects on the food web robustness than low-degree species
\citetext{\citealp{Dunne2002a}; \citealp[references
in][]{Cirtwill2018a}}. The trophic level offers information about a
species vertical position in the food web, and top predators at higher
trophic levels and primary producers at the base are particularly
important for food web dynamics through top-down and bottom-up control
\citep[references in][]{Cirtwill2018a}. Trophic similarity is an index
of trophic overlap considering the set of prey and predators for a pair
of species; it measures one of the most important aspects of a species'
niches, the trophic niche \citep{Martinez1991, Williams2000}.

Furthermore, we took species habitat affiliation into account, which
describes the physical position of a species within the ecosystem.
Species were categorised as: 1) benthic, if it lives on the seafloor; 2)
pelagic, if it lives close to the surface; 3) benthopelagic, if it moves
between and connects the aforementioned environments; 4) demersal, if it
lives and feeds on, or near, the bottom of the sea; and 5) land-based,
if the consumer is not strictly aquatic but feeds predominantly on
marine species. Species habitat affiliations were retrieved from
\citet{Jacob2011}.

To study the relationship between species mean IS (weighted property)
and the unweighted species properties, we performed linear regression
analyses between the log mean IS and each of the aforementioned
unweighted properties. Thus, we considered the IS as the dependent
variable and the given unweighted property as the independent variable,
and obtained the coefficients (slope and intercept) for the linear
model. Models were fitted using the least squares approach. We also
explored the mean IS distribution with the species' habitat.

@@

To analyse the impact of species on food web stability, we performed
extinction simulations deleting one species at a time, that is for every
extinction, network size was reduced by one species only. After each
extinction, we calculated the stability of the network minus the removed
species (489 nodes) and compared it with that of the whole network (490
nodes in total).

To estimate stability, we utilized the mean of the real part of the
maximum eigenvalue of the Jacobian (or community) matrix, employing
random values for the elements of the Jacobian while preserving the
predator-prey sign structure
\citep{Saravia2022, Allesina2008, Pawar2009}. The Jacobian matrix,
denoted as \(C\), is fundamental to understanding the population
dynamics and inter-species interactions within an ecological network.
Each element \(c_{ij}\) of \(C\) represents the effect of a change in
the \(j^{th}\) species' density on the \(i^{th}\) species, at
equilibrium. When \(i\) equals \(j\), \(c_{ij}\) reflects the dependence
of the \(i^{th}\) species on its own density, offering insights into
intra-species interactions and self-regulation. The maximum eigenvalue
describes the rate at which a small disturbance decays or amplifies over
time in the vicinity of an equilibrium. Precisely, the real parts of the
eigenvalues indicate the rate of exponential growth (if positive) or
decay (if negative) of perturbations. Thus, a more negative index is
indicative of a more stable food web with reduced probability of
perturbation amplification; in other words the network is more resilient
to disturbances.

We conducted 1000 simulations for the removal of each species,
calculating the maximum eigenvalue for the network in each case. For
each simulation, we compared this value against the median maximum
eigenvalue obtained from 1000 simulations of the complete network, thus
generating a distribution of differences. A positive difference
indicates that the food web's stability is greater without the targeted
species, suggesting that the species in question contributes to the
network's instability. Conversely, a negative difference implies that
the network is less stable without the species, indicating a stabilizing
effect. Due to the variability in the estimation of the eigenvalues, we
decided to consider that a substantial impact on stability was reached
when the proportion of either negative or positive differences within
this distribution exceeded 0.55, meaning that 55\% of the time the
difference in stability was positive or negative. A detailed description
of the stability calculations can be found in the Supplementary Material
(Figure S2).

To identify the species with the highest effect on food web stability
and their characteristics, we plotted the results of each species'
extinction and its effect on food web stability (i.e., the stability
difference between the whole network and the network minus one species)
against weighted (interaction strength), and the three unweighted
properties (trophic level, degree, and trophic similarity), and species
habitat affiliation.

@@

The statistical distribution that best fitted the empirical interaction
strength distribution of the Weddell Sea food web was a `log-Normal' due
to the high proportion of weak interactions and the existence of a few
strong interactions (Figure 3, Table S3).

Here it is important to note that we observed an important variability
when estimating each IS. The interactions that showed the greatest
variability, represented by a wide interquartile range, were those where
consumer and resource body masses were very small (approximately 1e-15
kg). The frequency distribution of the interquartile range is presented
in the Supplementary Material (Figure S1).
